Dismal Code: Studying the Evolution of Security Bugs

### 2013-04-08 docs/evol.tex
minor fix
minor corrections and ammendments

\begin{abstract}
%\boldmath
Security bugs are critical programming errors that can lead to serious
vulnerabilities in software. Such bugs may allow an attacker to take over
an application, steal data or prevent the application from working at all.
We used the Maven repository to study the
characteristics of security bugs individually and in relation to other software
bugs. In particular, we analyzed every project version of the repository by using a
popular static analysis tool. Then by taking advantage of
the repository's features (multiple versions for every project, dependencies) we studied
the evolution of security bugs through time, their persistence and their relationship with a) the
size of the corresponding version and b) other bug categories. Our findings
indicated that security bugs neither increase or decrease as a project evolves.
In addition, they are not eliminated in a way that is particularly
different from the other bugs. Also, their relation with the
size appeared to be different from the other categories. Finally, even if bugs
seem to flock together, severe security bugs
seem to be unassociated with other categories.
\end{abstract}

@@

In this work we:
\begin{itemize}
	\item Analyze how security bugs found through static analysis,
evolve over time. To achieve this, we inspect every project per version.
Our hypothesis is that security bugs should decrease as a project evolves since
they are critical and developers should eliminate them.
%We present evidence that defect counts remain relatively stable
%across project revisions.
	\item Examine security bug persistence between versions. 
We expect that security bugs should be eliminated earlier than other bugs.
%We find that a specific defect remains unfixed on average 
%for 3 revisions.
	\item Study the relation between security bugs and a project
version's size. Our hypothesis is that security bugs are proportional to a project
version's size. Here, size is defined in terms of bytecode.
%We show that many defect types are correlated with the project's
%size (with the notable exception of high-risk security defects).
	\item Examine the correlation of security bugs with other bug categories.
Our hypothesis is that security bugs flock together with bugs that
have to do with performance, coding practices and others.
%We show that a code base can have defect hotspots.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
  %\item Make available a carefully constructed dataset and a statistical toolkit for
   % performing analysis of the Java Maven ecosystem.
\end{itemize}

@@

A threat to the internal validity of our experiment could be the false alarms of the
FindBugs tool. The false positives and negatives of static analysis tools and
how they can be reduced, is an issue that has already been discussed in literature (see
Section~\ref{sec:rel}). Furthermore, consider a false positive alarm where an
security bug is reported, but it is invalidated based on the context of the
application (an {\sc sql} injection vulnerability~\cite{RL12} in an application that
requires no external input).

@@

In this work we analyzed a large software repository to examine the behavior
of security bugs. Specifically, we analyzed more than 260GB of interdependent
projcet versions to see how security bugs evolve over time, their
persistence, and their relation with other bug categories and the size in
terms of bytecode.

Our primary hypothesis was that security bugs and especially severe
ones, should be corrected as projects evolve. However, our findigs indicate that this is
not true. Contrary to our second research hypothesis, we showed that security
bugs are not eliminated in a way that is particularly different from the other bugs.
Also, having an average of two to three versions persistence in a sample where 60\%
of the projects have three versions, is not a positive result especially in the
case of the {\it Security High} bugs. Concerning the relation between severe security bugs and a project's size
we showed evidence that they are not proportionally related. Furthermore, the
pairwise correlations between all categories indicated that even though all the other
categories are related, severe bugs do not flock together with the other bugs.
Also, it is interesting to see that security bugs were one of the top two
bug Make available a carefully constructed dataset and a statistical toolkit for
    performing analysis of the Java Maven ecosystem.
categories existing in a large ecosystem. Finally, we highlighted the
domino effect, and showed evidence that indicates that the Linus' Law does not
apply in the case of the security bugs.

Contrary to the approaches that examine all versions
formed after every change that has been commited to the repository,
our observations are made from a different perspective.
The versions examined in this work were actual releases
of the projets. As a result we do not have an indication of how many changes
have been made between the releases.
In essence, these {\sc jar}s were the ones that were or still are,
out there in the wild, being used either as applications,
or dependencies of others.

>>>>

\begin{abstract}
%\boldmath
Security bugs are critical programming errors that can lead to serious
vulnerabilities in software. Such bugs may allow an attacker to take over
an application, steal data or prevent the application from working at all.
We used the projects stored in the Maven repository to study the
characteristics of security bugs individually and in relation to other software
bugs. In particular, we analyzed every project version of the repository by using a
popular static analysis tool. Then by taking advantage of
the repository's features (multiple versions for every project, dependencies) we studied
the evolution of security bugs through time, their persistence and their relationship with a) the
size of the corresponding version and b) other bug categories. Our findings
indicate that there is no simple relationship between the number of
security bugs and project evolution.
In addition, they are not eliminated in a way that is particularly
different from the other bugs. Their relation with
size appears to be different from the other categories. Finally, even if bugs
seem to flock together, severe security bugs
seem to be unassociated with other categories.
\end{abstract}

@@

In this work we:
\begin{itemize}
	\item Analyze how security bugs found through static analysis
evolve over time. To achieve this, we inspected every project per version.
Our hypothesis is that security bugs should decrease as a project evolves since
they are critical and developers should eliminate them.
%We present evidence that defect counts remain relatively stable
%across project revisions.
	\item Examine security bug persistence between versions. 
We expect that security bugs should be eliminated earlier than other bugs.
%We find that a specific defect remains unfixed on average 
%for 3 revisions.
	\item Study the relation between security bugs and a project
version's size. Our hypothesis is that security bugs are proportional to a project
version's size. Here, size is defined in terms of bytecode.
%We show that many defect types are correlated with the project's
%size (with the notable exception of high-risk security defects).
	\item Examine the correlation of security bugs with other bug categories.
Our hypothesis is that security bugs flock together with bugs that
have to do with performance, coding practices and others.
%We show that a code base can have defect hotspots.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
  %\item Make available a carefully constructed dataset and a statistical toolkit for
   % performing analysis of the Java Maven ecosystem.
\end{itemize}

@@

A threat to the internal validity of our experiment could be the false alarms of the
FindBugs tool. False positives and negatives of static analysis tools and
how they can be reduced is an issue that has already been discussed in literature (see
Section~\ref{sec:rel}). Furthermore, consider a false positive alarm where an
security bug is reported, but it is invalidated based on the context of the
application (an {\sc sql} injection vulnerability~\cite{RL12} in an application that
requires no external input).

@@

In this work we analyzed a large software repository to examine the
behavior of security bugs. Specifically, we analyzed more than 260GB
of interdependent project versions to see how security bugs evolve
over time, their persistence, their relation with other bug
categories, and their relationship with size in terms of bytecode.

Our primary hypothesis was that security bugs, and especially severe
ones, should be corrected as projects evolve. We found that, although
bugs do close over time in particular projects, we do not have an indication that across
projects they decrease as projects mature. Moreover, defect counts may
increase, as well as decrease in time. Contrary to our second research hypothesis, we showed that security
bugs are not eliminated in a way that is particularly different from the other bugs.
Also, having an average of two to three versions persistence in a sample where 60\%
of the projects have three versions, is not a positive result especially in the
case of the {\it Security High} bugs. Concerning the relation between severe security bugs and a project's size
we showed evidence that they are not proportionally related. Furthermore, the
pairwise correlations between all categories indicated that even though all the other
categories are related, severe bugs do not flock together with the other bugs.
Also, it is interesting to see that security bugs were one of the top two
bug Make available a carefully constructed dataset and a statistical toolkit for
    performing analysis of the Java Maven ecosystem.
categories existing in a large ecosystem. Finally, we highlighted the
domino effect, and showed evidence that indicates that the Linus' Law does not
apply in the case of the security bugs.

Contrary to the approaches that examine all versions
formed after every change that has been commited to the repository,
our observations are made from a different perspective.
The versions examined in this work were actual releases
of the projects. As a result we do not have an indication of how many changes
have been made between the releases.
In essence, these {\sc jar}s were the ones that were or still are,
out there in the wild, being used either as applications,
or dependencies of others.
